```{r, echo=FALSE}
data("bfi", package = "psychTools")
library(dplyr)
bfi <- mutate(bfi,
              gm = case_when(
                  age < 18 & gender == "male"   ~ "boy",
                  age < 18 & gender == "female" ~ "girl",
                  age >= 18 & gender == "male"  ~ "man",
                  age >= 18 & gender == "female"~ "woman",
                  TRUE ~ NA_character_),
              gm = as.factor(gm))
```


Question
========
Which commands correctly rename variables in the `bfi` dataset using `rename()` or `rename_with()` from `dplyr`?

```{r}
head(bfi)
```


Answerlist
----------
* `bfi <- rename(bfi, sex = gender)`
* `bfi <- rename_with(bfi, tolower)`
* `bfi <- rename_with(bfi, .fn = tolower, .cols = matches("\\d$"))`
* `bfi <- rename_with(bfi, toupper, .cols = 1:3)`
* `bfi <- rename(bfi, gendered_maturity = gm)`
* `bfi <- rename_with(bfi, .fn = toupper, .cols = starts_with("a"))`
* `bfi <- rename(bfi, gendered_maturity = gm, ed = education)`
* `bfi <- rename(bfi, gender = sex)`
* `bfi <- rename(bfi, "sex" <- "gender")`
* `rename(bfi, .fn = tolower, .cols = gender)`

Solution
========

Answerlist
----------
* TRUE. Renames one variable.
* TRUE. Renames all variables to lowercase.
* TRUE. Renames variables ending with a digit to lowercase.
* TRUE. Renames first three variables to uppercase.
* TRUE. Renames one variable.
* TRUE. Renames variables starting with "a" to uppercase.
* TRUE. Renames two variables.
* FALSE. Incorrect order of arguments.
* FALSE. Incorrect syntax for renaming.
* FALSE. Incorrect function usage.


Meta-information
================
exname: Renaming variables
extype: mchoice
exsolution: 1111111000
exshuffle: 5
