---
title: "Data Modification"
page-navigation: false
format:
  html: 
    toc: false
---

```{r, echo = FALSE, include = FALSE}
source(here::here("R", "buttons.R"))
```

Welcome to our module about data modification. Data rarely arrive in a perfect, ready-to-analyze form. Instead, they must be filtered, sorted, and transformed so that patterns become visible and variables take on meaningful structure. 

1. [Subsetting](/tutorials/Subsetting/index.qmd)
2. [Sorting](/tutorials/Sorting/index.qmd)
3. [Transforming data](/tutorials/Transformation/index.qmd)

Each tutorial will comprise a series of interactive lessons with practice problems embedded throughout.

## Learning Goals

After completing this module, you will be able to:

### Subsetting Data

1. Extract specific rows, columns, and elements from datasets using Base R and `dplyr`.
2. Apply logical conditions to filter data efficiently.
3. Use tools like `filter()` and `select()` to subset data clearly and reproducibly.

### Sorting Data

1. Reorder values or rows in ascending or descending order.
2. Use `dplyr::arrange()` to sort data frames by one or more variables.

### Transforming Data

1. Convert variables into factors with defined levels.
2. Create, modify, and standardize variables using `mutate()`.
3. Apply operations across multiple columns with `across()`.
4. Build new variables conditionally using `case_when()`.
5. Rename variables programmatically using `rename()` and `rename_with()`.

Click the button below to get started with the first tutorial.

```{r}
#| echo: false
#| results: asis
begin_button("/tutorials/Subsetting/index.html")
```
