---
title: "Module Template"
page-navigation: false
format:
  html: 
    toc: false
---

```{r, echo = FALSE, include = FALSE}
source(here::here("R", "buttons.R"))
```

Welcome to our template module. Every module comprises a series of topical *tutorials*, such as the three linked below.

1. [First Topic](/tutorials/first_tutorial/index.qmd)
1. [Second Topic](/tutorials/second_tutorial/index.qmd)

Each tutorial will comprise a series of interactive lessons with practice problems embedded throughout. Each tutorial
section should be wrapped up with a knowledge quiz.

## Learning Goals

After completing this module, you will be able to:

### Tutorial 1

1. A thing you should learn from Tutorial 1.
1. Another thing you should learn from Tutorial 1.
1. One more thing you should learn From Tutorial 1.

### Tutorial 2

1. A thing you should learn from Tutorial 2.
1. Another thing you should learn from Tutorial 2.

Click the button below to get started with the first tutorial.

```{r}
#| echo: false
#| results: asis
begin_button("/tutorials/first_tutorial/index.html")
```
