---
title: "Converting Variables into Factors"
---

```{r, echo=FALSE}
data("bfi", package = "psychTools")
```


A common transformation involves converting numeric or character variables into factors—categorical variables that have defined levels. Factors are essential for statistical modeling and for treating variables like gender, education level, or species as discrete categories rather than continuous quantities.

We can create a simple character vector and convert it into a factor using `as.factor()`:

```{r}
(animals <- sample(c("dog", "cat", "mongoose"), 25, TRUE))
(animalsF <- as.factor(animals))
levels(animalsF)
table(character = animals, factor = animalsF)
```

For character vectors, this quick-and-dirty approach works well. However, it’s less informative when applied to numeric variables because it treats numbers as labels rather than meaningful values:

```{r}
genderF <- as.factor(bfi$gender)
levels(genderF)
table(numeric = bfi$gender, factor = genderF)
```

To explicitly control the factor levels and their labels, use the `factor()` function:

```{r}
bfi0 <- bfi
bfi$gender <- factor(bfi$gender, labels = c("male", "female"))
levels(bfi$gender)
table(numeric = bfi0$gender, factor = bfi$gender)
```

This approach ensures that numeric codes are mapped to descriptive category names.










