---
title: "Logical subsetting"
---

```{r, echo=FALSE}
data(bfi, package = "psychTools")
```

```{webr-r}
#| autorun: true
#| context: setup
data(bfi, package = "psychTools")
```

Logical subsetting filters rows based on conditions that evaluate to `TRUE` or `FALSE`. Rows where the condition is `TRUE` are kept, and those where it is `FALSE` are removed. If a value in the condition is `NA`, it is treated as unknown and therefore excluded unless handled explicitly. This approach allows you to express complex filtering criteria in a direct and readable way.

We begin by creating a logical vector that identifies participants older than 30:

```{r}
(over30 <- bfi$age > 30)
```

This command returns a vector of `TRUE` and `FALSE` values. We can easily count how many participants are over 30 and what proportion they represent in the sample:

```{r}
sum(over30)
mean(over30)
```

We can now use this logical vector to subset the data frame, keeping only the rows where the condition is `TRUE`:

```{r}
bfi30 <- bfi[over30, ]
```

The new object `bfi30` contains only participants older than 30. We can check this by examining its dimensions and the minimum age value:

```{r}
dim(bfi30)
min(bfi30$age)
```

Logical conditions can be combined using the operators `&` (and), `|` (or), and `!` (not). For example, to select male participants (`gender == 1`) who are younger than 50 and have at least a bachelor’s degree (`education >= 4`), while excluding missing values for education, we can define a more complex filter:

```{r}
filter <- with(bfi, age < 50 & gender == 1 & !is.na(education) & education >= 4)
```

We can see how many cases meet these criteria:

```{r}
sum(filter)
```

Then we apply the filter to the dataset:

```{r}
bfi2 <- bfi[filter, ]
```

Finally, we check that the selection worked as expected:

```{r}
nrow(bfi2)
unique(bfi2$gender)
max(bfi2$age)
min(bfi2$education)
```

We can also build a quick cross-tabulation to see where the condition holds across values of a variable, which helps visually inspect the filtering logic:

```{r}
with(bfi, table(age, (gender == 1 & !is.na(education) & education >= 4)))
```


::: practice

## Practice

Use base R logic subsetting procedures to select the five neuroticism items for female minors out of the `bfi` data.

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}

```

## {{< var tab_title.hint >}}

Remember that if you want more information about the variables in the `bfi` dataset, you can check the documentation by running `?bfi` in the R console.

## {{< var tab_title.solution >}}

```{r}
bfi[bfi$gender == 2 & bfi$age < 18, c("N1", "N2", "N3", "N4", "N5")]
```
::::

:::




















